---
title: "make_stan_list"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{make_stan_list}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wztools)
library(magrittr)
```

For this vignette, we will use 2 attached synthetic datasets. First, for a single subject:

```{r}
single_subj_sim
```

This is a simulated dataset of accuracies and reaction times in two difficulty conditions (see `?single_subj_sim`). We will assume that conditions are blocked. We can use the `make_stan_list` function to transform the output to *stan*-appropriate format:

```{r}
var_names = c('choice', 'rt') #specifies which variables we want to summarise in multi-dimentional arrays
grouping_vars = c('cond', 'trial') #specifies by which dimentions will the output be grouped
grouping_vars_stan_names=c('numConds','numTrials')
stan_list = make_stan_list(dat = single_subj_sim,
                           var_names = var_names,
                           grouping_vars = grouping_vars,
                           grouping_vars_stan_names = grouping_vars_stan_names,
                           hierarchical = F # non-hierarchical setting is for a single subject
                           )
stan_list
```
The `stan_list` elements *rt* and *choice* are arrays, with *condition* and *trial* being the first and second dimention, respecively.

When using evidence accumulation to model choices, we might need to additionally add the minimal rt (rt bound), which we can do using the *ssm_prepro* argument:

```{r}
stan_list = make_stan_list(dat = single_subj_sim,
                           var_names = var_names,
                           grouping_vars = grouping_vars,
                           grouping_vars_stan_names = grouping_vars_stan_names,
                           hierarchical = F, # non-hierarchical setting is for a single subject
                           ssm_prepro=T)
stan_list
```
We can do this also for multisubject data:
```{r}
#pre-generated 10-subject dataset
sim
```

We just need to add subject identifier (`subjID`) as another dimension to our arrays. We can do this by adding `subjID` as the first grouping variable, and changing the `hierarchical` argument to `TRUE`:

```{r}
grouping_vars = c('subjID', 'cond', 'trial') 
grouping_vars_stan_names=c('numSubj', 'numConds','numTrials')

stan_list = make_stan_list(dat = sim,
                           var_names = var_names,
                           grouping_vars = grouping_vars,
                           grouping_vars_stan_names = grouping_vars_stan_names,
                           hierarchical = T, # non-hierarchical setting is for a single subject
                           ssm_prepro=T)
dim(stan_list$choice)
```
Now `choice` and `rt` are 3-dimentional arrays, with `subjID` being the first dimension. Such multi-dimensional arrays are a convenient way of inputting data into *Stan*.



